Generic Separable Algorithms for Distance Transformation: A Subquadratic Approach for Path Based Norms\thanks{This work has been mainly funded by ANR-11-BS02-009 and ANR-11-IDEX-0007-02 PALSE/2013/21 research grants.}

### 2014-02-13 latex/dgci2014/article.tex
ok

\begin{abstract}\sloppy
In many applications, separable algorithms have demonstrated their
efficiency to perform high performance volumetric computations, such
as distance transformation or medial axis extraction. In the
literature, several authors have discussed about the conditions on the
metric to be considered in a separable approach.  In this article, we
present generic separable algorithms to efficiently compute Voronoi
maps and distance transformations for a large class of
metrics. Focusing to path based norms (chamfer masks, neighborhood
sequences, ...), we detail a subquadratic algorithm to compute such
volumetric transformations. For example, we describe a
$O(\log^2{m}\cdot N^2)$ algorithm in dimension 2 for shapes in
$N\times N$ domain with chamfer norm with size $m$.

@@

\begin{table}
  \caption{Complexity summary.}
  \label{tab:final}
  \begin{center}
    \begin{tabular}{|c|c|c|c|c|}
      \hline
      Metric &\textsc{Closest}& \textsc{HiddenBy} & Sep. Voronoi Map & Reference\\
      \hline
      $L_2$ & $O(1)$ &  $O(1)$ & $\Theta(n\cdot N^n)$ & \cite{Hirata1996}\\
      $L_\infty$ & $O(1)$ & $O(1)$ &  $\Theta(n\cdot N^n)$ & \cite{roerdnik}\\
      $L_1$ & $O(1)$ &   $O(1)$ &$\Theta(n\cdot N^n)$& \cite{roerdnik}\\
      $L_p$  (exact pred.) & $O(\log{p})$ &
      $O(\log{p}\cdot\log{N})$ & $O(n\cdot
      N^n\cdot\log{p}\cdot\log{N})$& here and  \cite{dgtal}\\
      $L_p$  (inexact pred.) & $O(1)$ &
      $O(\log{N})$ & $O(n\cdot
      N^n\cdot\log{N})$& here and  \cite{dgtal}\\
      2D Chamfer norm &  $O(\log{m})$ &$O(\log^2{m})$
      &$O(\log^2{m}\cdot N^2)$& Theorem \ref{them}\\
      3D Chamfer norm &  $O(\log{m})$ &$O(\log^2{m})$
      &$O(\log^2{m}\cdot N^3)$& Theorem \ref{them2}\\
      nD Chamfer norm &  $O(n\cdot\log{f})$ &$O(n\cdot log^2{f}\cdot\log{N})$
      & $O(n^2\cdot N^n\cdot \log{f}\cdot\log{N})$& Corollary \ref{coro2}\\
      2D Neig. sequence norm  & $O(\log{f})$&
      $O(\log^2{f})$& $O(\log^2{}f\cdot N^2)$& Theorem \ref{them} and  \cite{DBLP:conf/dgci/NormandSE13}\\
      \hline
    \end{tabular}
  \end{center}
\end{table}

@@

In the literature
\cite{borgefors,Thiel_hdr,Strand2008,normandHDR,fouard:ivc:2005} many
authors have proposed algorithmic and/or analytic approaches to
construct Chamfer masks approximating the Euclidean metric and being
induced by norm. In the following, we focus on such \emph{chamfer
  norms}.  To evaluate distances between two digital points for a
given chamfer metric, direct formulations have been proposed with
simple geometrical interpretation:
\begin{definition}[Rational ball, minimal H-representation \cite{Thiel_hdr,DBLP:journals/pr/NormandE09}]
  Given a Chamfer mask $\M$ inducing a norm, the rational ball
  associated with $\M$ is the polytope
  \begin{equation}
\label{eq:4}
    \B = conv\left\{ \frac{\vec{v}_i}{w_i};\; (\vec{v}_i,w_i)\in\M \right \}\,.
  \end{equation}
Rational ball $\B$ can also be described in a H-representation of
polytope with minimal parameter \cite{DBLP:conf/dgci/NormandSE13}:
\begin{equation}
  P = \{ x\in\Z^n; Ax \leq y\;\}\quad \text{ such that }  \forall
  i\in[1\ldots f], \exists x\in P\quad A_ix=y_i\,.\footnote{$A_i$
    being the $i^{th}$ row of $A$.}
\end{equation}
$f$ being the number of row in $A$ and the number of facets in $\B$,   and is thus related to $|\M|$.
\end{definition}

An important result for distance computation can be formalized as follows:
\begin{proposition}[Direct Distance Computation \cite{DBLP:journals/pr/NormandE09}]
  Given a chamfer mask $\M$ inducing a norm and $(A,y)$ its minimal
  parameter H-representation, then for any $a\in\Z^n$, the chamfer
  distance of $a$ from the origin is
  \begin{equation}
    \label{eq:1}
    d_\M(O,a) =  \max_{1\leq i \leq l} \{ A_i a^T\}\,.
  \end{equation}
\end{proposition}

@@

\begin{algorithm}[H]\footnotesize
  \KwData{Input binary map $I_X$ if $i=1$ or  partial Voronoi map $\Pi_X$ obtained for
    dimension $1<j<i$, and a 1D segment $S$ with points
    $\{s_1,\ldots,s_{N_i}\}$ sorted by their $i^{th}$ coordinate.}
  \KwResult{Updated partial Voronoi map $\Pi_X$ along $S$.}
\If{ i = 1}
{ $k=0$\;
\ForEach{point $s_i$ in $S$}{
    \If{$I_X(s_i) = 1$}
     {
       $L_S[k] = s_i$\;
       $k++$\;
     }
}
}
\Else{
 ${L}_S[0]=\Pi_X(s_1)$\;
  ${L}_S[1]=\Pi_X(s_2)$\;
  $k=2$ , $l=3$\;
  \While{$l\leq N_i$}{
    $w = \Pi_X(s_l)$\;
    \While{$k\geq 2$ and \textsc{HiddenBy}(${L}_S[k-1],{L}_S[k],w, S$)}
          {
         $ k--$\;
        }
        $k++$ , $l++$\;
        ${L}_S[k]=w$\;
  }
}
  \ForEach{point  $s$ in $S$ by increasing $i^{th}$ coordinate}
  {
    \While{($k < |L_S|$) and not(\textsc{Closest}($s$, $L_S[k]$, $L_S[k+1]$))}
          {
            $k++$\;
          }
    $\Pi_X[s] = L_S[k]$\;
  }

  \caption{Voronoi map pruning along 1D segment $S$ along the $i^{th}$
    dimension.\label{alg:voro}}
\end{algorithm}

@@

Algorithm \ref{alg:voro} describe the 1D processes to perform on each
row, columns and higher dimension spans. In this process, metric
information are embedded in the following two predicates (see Fig.~\ref{fig:predicates}):
\textsc{Closest}$(a, b, c)$: given three points $a,b,c\in\Z^n$ this
predicate returns true if $d(a,b) < d(a,c)$. \textsc{HiddenBy}$(a, b,
c, S)$: given three points $a,b,c\in\Z^n$ such that $a_i<b_i<c_i$ and
a 1D image span $S^i$, this predicates returns true if there is no
$s\in S$ such that
    \begin{equation}
      d(b,s) < d(a,s)\text{ and }  d(b,s) < d(c,s)\,.
    \end{equation}

\begin{figure}
  \begin{center}
    \subfigure[]{ \begin{overpic}[width=3cm]{figs/hiddenByTrue}
    \put(32,5){$a$}
    \put(5,35){$b$}
    \put(25,75){$c$}
    \put(60,-10){$S$}
    \end{overpic}}
    \subfigure[]{\begin{overpic}[width=3cm]{figs/hiddenByFalse}
    \put(37,8){$a$}
    \put(23,37){$b$}
    \put(35,72){$c$}
    \put(60,-10){$S$}
      \end{overpic}}~~~~~~~~~~
    \subfigure[]{\begin{overpic}[width=1.68cm]{figs/closest}
    \put(-5,52){$a$}
    \put(2,78){$b$}
    \put(47,55){$c$}
    \put(38,-10){$S$}
      \end{overpic}
    }
  \end{center}
  \caption{xxxxxxxxxxx}
  \label{fig:predicates}
\end{figure}

@@

@@

Let us first suppose that we don't have $e$ but we know the wedges
$(v_{k},v_{k+1})$ (resp. $(v_j,v_{j+1})$) such that the vector
$(e-a)^T$ (resp. $(e-b)^T$) should belong to (see Fig.XXXXX). In this
situation, we know that $e$ is the solution of
\begin{equation}
\label{eq:3}
  A_k\cdot  (e-a)^T = A_j\cdot (e -b)^T\;.
\end{equation}
(since $e\in S$, we have one linear equation for one unknown
$e_i$). As a consequence, if we know the wedges the Voronoi edge
belongs to, we have the abscissa in $O(1)$ (see Algorithm \ref{alg:chamedge}).
\begin{algorithm}[h]\footnotesize
  $(v_k,v_{k+1})$ = \textsc{VoronoiEdgeWedge}$(a,b,\vec{v}_1,\vec{v}_m, S)$\;
  $(v_j,v_{j+1})$ = \textsc{VoronoiEdgeWedge}$(b,a,\vec{v}_1,\vec{v}_m, S)$\;

@@

The dichotomic search shrinks the set of vectors
$\{\vec{v}_i,\ldots,\vec{v}_j\}$ to end up is a wedge $(v_k,v_{k+1})$
such that the intersection point between the straight line $(a +
\vec{v}_k)$ and $S$ is in the Voronoi cell of $a$ and such that the
intersection between $(a + \vec{v}_k)$ and $S$ is in the Voronoi cell
of $b$. Algorithm~\ref{alg:chamedgew} thus first compute the
intersection points associated with a wedge $(\vec{v}_{i+(j-i)/2},
\vec{v}_{i+(j-i)/2+1})$ (lines $5-6$); Evaluates the distances at
these points (lines $7-8$) and then decides which set
$\{\vec{v}_i,\ldots,\vec{v}_{i+(j-i)/2}\}$ or
$\{\vec{v}_{i+(j-i)/2},\ldots,\vec{v}_j\}$ has to be considered for
the recursive call (lines $14-25$).

@@

\subsection{Subquadratic Algorithms in Higher Dimension}
\label{sec:subq-algor-high}
In this section, we consider a general chamfer norm in arbitrary
dimension $n$.  First, let us discuss about chamfer mask
combinatorics. If $m$ denotes the number of weighted vectors of $\M$,
its rational ball $\B$ has $O(m^{\lfloor \frac{n}{2}\rfloor})$
$i-$facets ($0 \leq i \leq d$) \cite{deberg}. If we denote by $f$ the
number of $(n-1)-$facets of $\B$ (\emph{i.e.} number of row in the
H-representation of $\B$), we have by duality the result that $|\B| =
O(f^{\lfloor \frac{n}{2}\rfloor})$. As a consequence, we have the
following distance evaluation result:

\begin{lemma}
\label{lem:distnD}
   Let $\M$ be a chamfer norm whose rational ball $\B$ has $f$
   $(n-1)-$facets in dimension $n$, then distance computation and thus
   \textsc{Closest} predicate are in (amortized) $O(n\cdot\log{f})$ with
   $O\left (\frac{f^{\lfloor \frac{n}{2}\rfloor}}{(\log{ f})^{\lfloor
       \frac{n}{2}\rfloor-\delta}}\right)$ space and preprocessing
   time\footnote{$\delta$ is arbitrarily small positive constant.}.
\end{lemma}
\begin{proof}
 Similarly to the 2D case, the distance $d_\M(O,a)$ for $a\in\Z^n$ is
 given by first solving a ray-shooting problem on convex polytopes
 which consists in first computing the $(n-1)$-facet of $\B$ pierced by the
 ray $(O,a)$. Once the facet is obtained, the associated $A_k$ row is
 used to evaluate
   $d_\M(O,a) = A_k\cdot a^T$
in $O(n)$.  Following \cite{DBLP:journals/dcg/MatousekS93}, such
ray-shooting requests on convex polytopes can be solved in
$O(\log{f})$ thanks to a preprocessing in $O\left (\frac{f^{\lfloor
    \frac{n}{2}\rfloor}}{(\log{ f})^{\lfloor
    \frac{n}{2}\rfloor-\delta}}\right)$. Please note that this
preprocessing time is roughly equivalent to the convex hull
computation in higher dimension which is in $O(f^{\lfloor
  \frac{n}{2}\rfloor})$. Hence, preprocessing for ray-shooting can be
done while computing the $\B$ rational ball using
Eq.~(\ref{eq:4}).$\Box$
\end{proof}

From Lemmas~\ref{lem:generic} and~\ref{lem:distnD}, we first have:
\begin{corollary}
\label{coro2}
   Let $\M$ be a chamfer norm whose rational ball $\B$ has $f$
   $(n-1)-$facets in dimension $n$, separable exact Voronoi Map
   $\Pi_X$ can be obtained in     $O(n^2\cdot N^n\cdot
   \log{f}\cdot\log{N})$,
thanks to a preprocessing in $O\left (\frac{f^{\lfloor
    \frac{n}{2}\rfloor}}{(\log{ f})^{\lfloor
    \frac{n}{2}\rfloor-\delta}}\right)$.
\end{corollary}

In higher dimension, we can still expect fast
\textsc{VoronoiEdgeWedge} function. Let us consider the 3D case
$n=3$ (and thus $f=O(m)$). In this situation, it turns out that the problem can be turned
into a simple, 2D one: Given two points $a,b\in\Z^3$ and an isothetic
line $l$, we need to compute the intersection between the Voronoi
$2-$facet defined by $a$ and $b$ and $l$. Let us suppose that this the
intersection point exists and that $a$, $b$ and $l$ are in
\emph{general position} ($a\not\subset l$, $b\not\subset l$).  Then,
the intersection point belongs to the two Euclidean planes $P_a$ and
$P_b$, $P_a$ being respectively the Euclidean plane containing $a$ and
$l$ (resp. $P_b$ defined by $b$ and $l$), see Fig.\ref{fig:3D}.

To construct a variant of \textsc{VoronoiEdgeWedge} in 3D, we just
have to replace the line 4 of Algorithm~\ref{alg:chamedgew} by the
following steps: First, we construct $\vec{v} = (\vec{v}_i +
\vec{v}_j)/2$. If we suppose that $\vec{v}_i$ and $\vec{v}_j$ belongs
to $P_a$, $\vec{v}$ also belongs to $P_a$. Then, we perform a 3D
ray-shooting to get the facet $F$ of $\B$ pierced by vector
$\vec{v}$. From this facet $F$, we construct the 2D
$(\vec{v}_{mid},\vec{v}_{mid+1})$ wedge on $P_a$ defined by the
intersection of $F$ and $P_a$. Once $(\vec{v}_{mid},\vec{v}_{mid+1})$
is obtained, we proceed with the rest of the \textsc{VoronoiEdgeWedge}
function (point intersection construction, distance evaluations and
recursive calls). Please note that the 3D ray shooting can be obtained
in $O(\log{m})$. Hence, we have:

\begin{lemma}
\label{lem:wedge3D}
  Let $\M$ be a chamfer norm with $m$ weighted vectors in dimension
  3. We assume that the $m$ vectors span uniformly the set of
  directions in $\R^3$, the 3D version of \textsc{VoronoiEdgeWedge} is
  in
  \begin{equation}
    O(\log^2{m})\,.
  \end{equation}
\end{lemma}

\begin{proof}
  As in dimension 2, each step can be done in $O(\log{m})$
  computations. Please note that ray-shooting, and intersection
  computation between a (rational slope) facet $F$ and the (rational)
  plane $P_a$ can be obtained in $O(1)$.  If we denote by $\M_{P_a}$
  be the 2D chamfer norm induced by $\B\cap P_a$. The hypothesis that
  the vectors uniformly span the sphere of directions is here to
  ensure that if we split $\M_{P_a}$ along the mid direction
  $\vec{v} =(\vec{v}_i + \vec{v}_j)/2$, both subsets of vectors have the same
  size. If so, $O(\log{m})$ recursive calls is required. The stop
  criterion of the recurrence could be that $\vec{v}$ belong to the
  same facet as both $\vec{v}_i$ and $\vec{v}_j$.$\Box$
\end{proof}

From Lemma~\ref{lem:wedge3D} and proof of Theorem~\ref{them}, we
conclude with the following result:

\begin{theorem}
\label{them2}
  Let $\M$ be a chamfer norm with $m$ weighted vectors in dimension 3,
  the Voronoi map $\Pi_X$ and thus the distance transformation of $X$
  for metric $d_{M}$ can be obtained in
  \begin{equation}
    O(\log^2{m}\cdot N^3)\;.
  \end{equation}
\end{theorem}

>>>>

\begin{abstract}\sloppy
In many applications, separable algorithms have demonstrated their
efficiency to perform high performance volumetric computations, such
as distance transformation or medial axis extraction. In the
literature, several authors have discussed about the conditions on the
metric to be considered in a separable approach.  In this article, we
present generic separable algorithms to efficiently compute Voronoi
maps and distance transformations for a large class of
metrics. Focusing to path based norms (chamfer masks, neighborhood
sequences, ...), we detail a subquadratic algorithm to compute such
volumetric transformations in dimension 2. More precisely, we describe
a $O(\log^2{m}\cdot N^2)$ algorithm in dimension 2 for shapes in
$N\times N$ domain with chamfer norm with size $m$.

@@

\begin{table}
  \caption{Complexity summary.}
  \label{tab:final}
  \begin{center}
    \begin{tabular}{|c|c|c|c|c|}
      \hline
      Metric &\textsc{Closest}& \textsc{HiddenBy} & Sep. Voronoi Map & Reference\\
      \hline
      $L_2$ & $O(1)$ &  $O(1)$ & $\Theta(n\cdot N^n)$ & \cite{Hirata1996}\\
      $L_\infty$ & $O(1)$ & $O(1)$ &  $\Theta(n\cdot N^n)$ & \cite{roerdnik}\\
      $L_1$ & $O(1)$ &   $O(1)$ &$\Theta(n\cdot N^n)$& \cite{roerdnik}\\
      $L_p$  (exact pred.) & $O(\log{p})$ &
      $O(\log{p}\cdot\log{N})$ & $O(n\cdot
      N^n\cdot\log{p}\cdot\log{N})$& here and  \cite{dgtal}\\
      $L_p$  (inexact pred.) & $O(1)$ &
      $O(\log{N})$ & $O(n\cdot
      N^n\cdot\log{N})$& here and  \cite{dgtal}\\
      2D Chamfer norm &  $O(\log{m})$ &$O(\log^2{m})$
      &$O(\log^2{m}\cdot N^2)$& Theorem \ref{them}\\
  %    3D Chamfer norm &  $O(\log{m})$ &$O(\log^2{m})$
  %    &$O(\log^2{m}\cdot N^3)$& Theorem \ref{them2}\\
  %    nD Chamfer norm &  $O(n\cdot\log{f})$ &$O(n\cdot log^2{f}\cdot\log{N})$
  %    & $O(n^2\cdot N^n\cdot \log{f}\cdot\log{N})$& Corollary \ref{coro2}\\
      2D Neig. sequence norm  & $O(\log{f})$&
      $O(\log^2{f})$& $O(\log^2{}f\cdot N^2)$& Th.~\ref{them} and  \cite{DBLP:conf/dgci/NormandSE13}\\
      \hline
    \end{tabular}
  \end{center}
\end{table}

@@

In the literature
\cite{borgefors,Thiel_hdr,Strand2008,normandHDR,fouard:ivc:2005} many
authors have proposed algorithmic and/or analytic approaches to
construct Chamfer masks approximating the Euclidean metric and being
induced by norm. In the following, we focus on such \emph{chamfer
  norms}.  To evaluate distances between two digital points for a
given chamfer metric, direct formulations have been proposed with
simple geometrical interpretation:
\begin{definition}[Rational ball, minimal H-representation \cite{Thiel_hdr,DBLP:journals/pr/NormandE09}]
  Given a Chamfer mask $\M$ inducing a norm, the rational ball
  associated with $\M$ is the polytope
  \begin{equation}
\label{eq:4}
    \B = conv\left\{ \frac{\vec{v}_i}{w_i};\; (\vec{v}_i,w_i)\in\M \right \}\,.
  \end{equation}
Rational ball $\B$ can also be described in a H-representation of
polytope with minimal parameter \cite{DBLP:conf/dgci/NormandSE13}:
\begin{equation}
  P = \{ x\in\Z^n; Ax \leq y\;\}\quad \text{ such that }  \forall
  i\in[1\ldots f], \exists x\in P\quad A_ix=y_i\,.\footnote{$A_i$
    being the $i^{th}$ row of $A$.}
\end{equation}
$f$ being the number of row in $A$ and the number of facets in $\B$,   and is thus related to $|\M|$.
\end{definition}
An important result for distance computation can be formalized as follows:
\begin{proposition}[Direct Distance Computation \cite{DBLP:journals/pr/NormandE09}]
  Given a chamfer mask $\M$ inducing a norm and $(A,y)$ its minimal
  parameter H-representation, then for any $a\in\Z^n$, the chamfer
  distance of $a$ from the origin is
  \begin{equation}
    \label{eq:1}
    d_\M(O,a) =  \max_{1\leq i \leq l} \{ A_i a^T\}\,.
  \end{equation}
\end{proposition}

@@

@@

Algorithm \ref{alg:voro} describe the 1D processes to perform on each
row, columns and higher dimension spans. In this process, metric
information are embedded in the following two predicates (see Fig.~\ref{fig:predicates}):
\textsc{Closest}$(a, b, c)$: given three points $a,b,c\in\Z^n$ this
predicate returns true if $d(a,b) < d(a,c)$. \textsc{HiddenBy}$(a, b,
c, S)$: given three points $a,b,c\in\Z^n$ such that $a_i<b_i<c_i$ and
a 1D image span $S^i$, this predicates returns true if there is no
$s\in S$ such that
    \begin{equation}
      d(b,s) < d(a,s)\text{ and }  d(b,s) < d(c,s)\,.
    \end{equation}
\begin{algorithm}[H]\footnotesize
  \KwData{Input binary map $I_X$ if $i=1$ or  partial Voronoi map $\Pi_X$ obtained for
    dimension $1<j<i$, and a 1D segment $S$ with points
    $\{s_1,\ldots,s_{N_i}\}$ sorted by their $i^{th}$ coordinate.}
  \KwResult{Updated partial Voronoi map $\Pi_X$ along $S$.}
\If{ i = 1}
{ $k=0$\;
\ForEach{point $s_i$ in $S$}{
    \If{$I_X(s_i) = 1$}
     {
       $L_S[k] = s_i$\;
       $k++$\;
     }
}
}
\Else{
 ${L}_S[0]=\Pi_X(s_1)$\;
  ${L}_S[1]=\Pi_X(s_2)$\;
  $k=2$ , $l=3$\;
  \While{$l\leq N_i$}{
    $w = \Pi_X(s_l)$\;
    \While{$k\geq 2$ and \textsc{HiddenBy}(${L}_S[k-1],{L}_S[k],w, S$)}
          {
         $ k--$\;
        }
        $k++$ , $l++$\;
        ${L}_S[k]=w$\;
  }
}
  \ForEach{point  $s$ in $S$ by increasing $i^{th}$ coordinate}
  {
    \While{($k < |L_S|$) and not(\textsc{Closest}($s$, $L_S[k]$, $L_S[k+1]$))}
          {
            $k++$\;
          }
    $\Pi_X[s] = L_S[k]$\;
  }

  \caption{Voronoi map pruning along 1D segment $S$ along the $i^{th}$
    dimension.\label{alg:voro}}
\end{algorithm}
\begin{figure}
  \begin{center}
    \subfigure[]{ \begin{overpic}[width=2cm]{figs/hiddenByTrue}
    \put(26,5){$a$}
    \put(0,35){$b$}
    \put(25,75){$c$}
    \put(55,-15){$S$}
    \end{overpic}}
    \subfigure[]{\begin{overpic}[width=2cm]{figs/hiddenByFalse}
    \put(37,8){$a$}
    \put(23,37){$b$}
    \put(30,72){$c$}
    \put(55,-15){$S$}
    \end{overpic}}~~~~~~~~~~
    \subfigure[]{\begin{overpic}[width=1.1cm]{figs/closest}
    \put(-5,52){$a$}
    \put(2,78){$b$}
    \put(47,55){$c$}
    \put(38,-15){$S$}
      \end{overpic}
    }
  \end{center}
  \caption{xxxxxxxxxxx}
  \label{fig:predicates}
\end{figure}

@@

\begin{figure}
 {\includegraphics[width=1.65cm]{image-ball-1}}
 {\includegraphics[width=1.65cm]{image-ball-1_3999999999999999.png}}
 {\includegraphics[width=1.65cm]{image-ball-2}}
 {\includegraphics[width=1.65cm]{image-ball-4}}
 {\includegraphics[width=1.65cm]{image-ball-43_100000000000001.png}}
 {\includegraphics[width=1.65cm]{image-ball-chamf34}}
 {\includegraphics[width=1.65cm]{image-ball-chamf5711}}
  \caption{Balls for different metrics satisfying Proposition \ref{prop:voronoi-map-from-1}: \emph{(from
      left to right)} $L_1$, $L_{1.4}$, $L_2$, $L_4$, $L_{43.0}$,
    $\M_{3-4}$ and $\M_{5-7-11}$.}
  \label{fig:metrics}
\end{figure}

@@

Let us first suppose that we don't have $e$ but we know the wedges
$(v_{k},v_{k+1})$ (resp. $(v_j,v_{j+1})$) such that the vector
$(e-a)^T$ (resp. $(e-b)^T$) should belong to (see Fig.~\ref{fig:algo}). In this
situation, we know that $e$ is the solution of
\begin{equation}
\label{eq:3}
  A_k\cdot  (e-a)^T = A_j\cdot (e -b)^T\;.
\end{equation}
(since $e\in S$, we have one linear equation for one unknown
$e_i$). As a consequence, if we know the wedges the Voronoi edge
belongs to, we have the abscissa in $O(1)$ (see Algorithm \ref{alg:chamedge}).
\begin{algorithm}[h]\footnotesize
  $(v_k,v_{k+1})$ = \textsc{VoronoiEdgeWedge}$(a,b,\vec{v}_1,\vec{v}_m, S)$\;
  $(v_j,v_{j+1})$ = \textsc{VoronoiEdgeWedge}$(b,a,\vec{v}_1,\vec{v}_m, S)$\;

@@

\begin{figure}[htbp]
  \begin{center}
    \subfigure[]{
      \begin{overpic}[width=3.2cm]{init}
        \put(-5,60){$a$}
        \put(85,40){$b$}
        \put(42,105){$S$}
      \end{overpic}
    }~~~~~~~~~
    \subfigure[]{
      \begin{overpic}[width=3.2cm]{algoMid}
      \put(-5,60){$a$}
        \put(85,40){$b$}
        \put(42,105){$S$}
        \put(12,60){$\vec{v_{mid}}$}
        \put(13,80){$\vec{v_{mid+1}}$}
        \put(-5,98){$\vec{v_j}$}
        \put(5,35){$\vec{v_i}$}
        \put(50,90){$p_{mid+1}$}
        \put(50,58){$p_{mid}$}
      \end{overpic}
    }~~~~~~~~~
    \subfigure[]{
      \begin{overpic}[width=3.2cm]{algoEnd}
      \put(-5,60){$a$}
        \put(85,40){$b$}
        \put(42,105){$S$}
        \put(50,50){$e$}
        \put(10,50){$\vec{v_{k}}$}
        \put(11,70){$\vec{v_{k+1}}$}
        \put(60,40){$\vec{v_j}$}
        \put(67,57){$\vec{v_{j+1}}$}
      \end{overpic}
    }
  \end{center}
  \caption{Illustrations notations of \textsc{VoronoiEdgeWedge} and
    \textsc{VoronoiEdge}: $(a)$ initial problem, given two sites $a$
    and $b$, we want to compute the intersection between $S$ and the
    Voronoi edge of $a$ and $b$. $(b)$ an internal step of
    \textsc{VoronoiEdgeWedge} to reduce the set of directions of $\M$
    at $a$ (here the next recursive call will be on
    $(v_i,v_{mid})$).$(c)$ final step of \textsc{VoronoiEdge} where
    both wedges have been obtained and thus $e$ can be computed.}
  \label{fig:algo}
\end{figure}

The dichotomic search shrinks the set of vectors
$\{\vec{v}_i,\ldots,\vec{v}_j\}$ to end up is a wedge $(v_k,v_{k+1})$
such that the intersection point between the straight line $(a +
\vec{v}_k)$ and $S$ is in the Voronoi cell of $a$ and such that the
intersection between $(a + \vec{v}_k)$ and $S$ is in the Voronoi cell
of $b$ (see Fig.~\ref{fig:algo}$-(b)$). Algorithm~\ref{alg:chamedgew} thus first compute the
intersection points associated with a wedge $(\vec{v}_{i+(j-i)/2},
\vec{v}_{i+(j-i)/2+1})$ (lines $5-6$); Evaluates the distances at
these points (lines $7-8$) and then decides which set
$\{\vec{v}_i,\ldots,\vec{v}_{i+(j-i)/2}\}$ or
$\{\vec{v}_{i+(j-i)/2},\ldots,\vec{v}_j\}$ has to be considered for
the recursive call (lines $14-25$).

@@

%% \subsection{Subquadratic Algorithms in Higher Dimension}
%% \label{sec:subq-algor-high}
%% In this section, we consider a general chamfer norm in arbitrary
%% dimension $n$.  First, let us discuss about chamfer mask
%% combinatorics. If $m$ denotes the number of weighted vectors of $\M$,
%% its rational ball $\B$ has $O(m^{\lfloor \frac{n}{2}\rfloor})$
%% $i-$facets ($0 \leq i \leq d$) \cite{deberg}. If we denote by $f$ the
%% number of $(n-1)-$facets of $\B$ (\emph{i.e.} number of row in the
%% H-representation of $\B$), we have by duality the result that $|\B| =
%% O(f^{\lfloor \frac{n}{2}\rfloor})$. As a consequence, we have the
%% following distance evaluation result:

%% \begin{lemma}
%% \label{lem:distnD}
%%    Let $\M$ be a chamfer norm whose rational ball $\B$ has $f$
%%    $(n-1)-$facets in dimension $n$, then distance computation and thus
%%    \textsc{Closest} predicate are in (amortized) $O(n\cdot\log{f})$ with
%%    $O\left (\frac{f^{\lfloor \frac{n}{2}\rfloor}}{(\log{ f})^{\lfloor
%%        \frac{n}{2}\rfloor-\delta}}\right)$ space and preprocessing
%%    time\footnote{$\delta$ is arbitrarily small positive constant.}.
%% \end{lemma}
%% \begin{proof}
%%  Similarly to the 2D case, the distance $d_\M(O,a)$ for $a\in\Z^n$ is
%%  given by first solving a ray-shooting problem on convex polytopes
%%  which consists in first computing the $(n-1)$-facet of $\B$ pierced by the
%%  ray $(O,a)$. Once the facet is obtained, the associated $A_k$ row is
%%  used to evaluate
%%    $d_\M(O,a) = A_k\cdot a^T$
%% in $O(n)$.  Following \cite{DBLP:journals/dcg/MatousekS93}, such
%% ray-shooting requests on convex polytopes can be solved in
%% $O(\log{f})$ thanks to a preprocessing in $O\left (\frac{f^{\lfloor
%%     \frac{n}{2}\rfloor}}{(\log{ f})^{\lfloor
%%     \frac{n}{2}\rfloor-\delta}}\right)$. Please note that this
%% preprocessing time is roughly equivalent to the convex hull
%% computation in higher dimension which is in $O(f^{\lfloor
%%   \frac{n}{2}\rfloor})$. Hence, preprocessing for ray-shooting can be
%% done while computing the $\B$ rational ball using
%% Eq.~(\ref{eq:4}).$\Box$
%% \end{proof}

%% From Lemmas~\ref{lem:generic} and~\ref{lem:distnD}, we first have:
%% \begin{corollary}
%% \label{coro2}
%%    Let $\M$ be a chamfer norm whose rational ball $\B$ has $f$
%%    $(n-1)-$facets in dimension $n$, separable exact Voronoi Map
%%    $\Pi_X$ can be obtained in     $O(n^2\cdot N^n\cdot
%%    \log{f}\cdot\log{N})$,
%% thanks to a preprocessing in $O\left (\frac{f^{\lfloor
%%     \frac{n}{2}\rfloor}}{(\log{ f})^{\lfloor
%%     \frac{n}{2}\rfloor-\delta}}\right)$.
%% \end{corollary}

%% In higher dimension, we can still expect fast
%% \textsc{VoronoiEdgeWedge} function. Let us consider the 3D case
%% $n=3$ (and thus $f=O(m)$). In this situation, it turns out that the problem can be turned
%% into a simple, 2D one: Given two points $a,b\in\Z^3$ and an isothetic
%% line $l$, we need to compute the intersection between the Voronoi
%% $2-$facet defined by $a$ and $b$ and $l$. Let us suppose that this the
%% intersection point exists and that $a$, $b$ and $l$ are in
%% \emph{general position} ($a\not\subset l$, $b\not\subset l$).  Then,
%% the intersection point belongs to the two Euclidean planes $P_a$ and
%% $P_b$, $P_a$ being respectively the Euclidean plane containing $a$ and
%% $l$ (resp. $P_b$ defined by $b$ and $l$), see Fig.\ref{fig:3D}.

%% To construct a variant of \textsc{VoronoiEdgeWedge} in 3D, we just
%% have to replace the line 4 of Algorithm~\ref{alg:chamedgew} by the
%% following steps: First, we construct $\vec{v} = (\vec{v}_i +
%% \vec{v}_j)/2$. If we suppose that $\vec{v}_i$ and $\vec{v}_j$ belongs
%% to $P_a$, $\vec{v}$ also belongs to $P_a$. Then, we perform a 3D
%% ray-shooting to get the facet $F$ of $\B$ pierced by vector
%% $\vec{v}$. From this facet $F$, we construct the 2D
%% $(\vec{v}_{mid},\vec{v}_{mid+1})$ wedge on $P_a$ defined by the
%% intersection of $F$ and $P_a$. Once $(\vec{v}_{mid},\vec{v}_{mid+1})$
%% is obtained, we proceed with the rest of the \textsc{VoronoiEdgeWedge}
%% function (point intersection construction, distance evaluations and
%% recursive calls). Please note that the 3D ray shooting can be obtained
%% in $O(\log{m})$. Hence, we have:

%% \begin{lemma}
%% \label{lem:wedge3D}
%%   Let $\M$ be a chamfer norm with $m$ weighted vectors in dimension
%%   3. We assume that the $m$ vectors span uniformly the set of
%%   directions in $\R^3$, the 3D version of \textsc{VoronoiEdgeWedge} is
%%   in
%%   \begin{equation}
%%     O(\log^2{m})\,.
%%   \end{equation}
%% \end{lemma}

%% \begin{proof}
%%   As in dimension 2, each step can be done in $O(\log{m})$
%%   computations. Please note that ray-shooting, and intersection
%%   computation between a (rational slope) facet $F$ and the (rational)
%%   plane $P_a$ can be obtained in $O(1)$.  If we denote by $\M_{P_a}$
%%   be the 2D chamfer norm induced by $\B\cap P_a$. The hypothesis that
%%   the vectors uniformly span the sphere of directions is here to
%%   ensure that if we split $\M_{P_a}$ along the mid direction
%%   $\vec{v} =(\vec{v}_i + \vec{v}_j)/2$, both subsets of vectors have the same
%%   size. If so, $O(\log{m})$ recursive calls is required. The stop
%%   criterion of the recurrence could be that $\vec{v}$ belong to the
%%   same facet as both $\vec{v}_i$ and $\vec{v}_j$.$\Box$
%% \end{proof}

%% From Lemma~\ref{lem:wedge3D} and proof of Theorem~\ref{them}, we
%% conclude with the following result:

%% \begin{theorem}
%% \label{them2}
%%   Let $\M$ be a chamfer norm with $m$ weighted vectors in dimension 3,
%%   the Voronoi map $\Pi_X$ and thus the distance transformation of $X$
%%   for metric $d_{M}$ can be obtained in
%%   \begin{equation}
%%     O(\log^2{m}\cdot N^3)\;.
%%   \end{equation}
%% \end{theorem}
